Simulating DeLosh 1997

### 2023-11-05 Simulations/DeLosh97_Sim.qmd
update DeLosh simulation functions

```{r}
#lapply(c('tidyverse','data.table','igraph','ggraph','kableExtra'),library,character.only=TRUE))
pacman::p_load(tidyverse,data.table,igraph,ggraph,kableExtra)

@@

ggplot() + geom_rect(data = odf,
            mapping = aes(xmin = xmin, ymin = ymin, 
                          xmax = xmax, ymax = ymax, 
                          fill = type, colour = type),alpha = 0.5) +
  geom_text(data=odf,aes(x=x,y=y,label=label,size=3)) +
  geom_path(data=plot_edges,mapping=aes(x=x,y=y,group=id,alpha=weight)) +
  # geom_rect(aes(xmin=-1.05,xmax=-.95,ymin=-10,ymax=5),color="red",alpha=.1)+
  # geom_rect(aes(xmin=-0.05,xmax=.05,ymin=-10,ymax=5),color="blue",alpha=.1) +
  theme_void()

@@

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
alm.response <- function(input=1,c) {
input.activation <- exp(-c*(input.layer - input)^2)
input.activation <<- input.activation/sum(input.activation)
#print(length(input.activation)); print(dim(weight.mat))
output.activation <<- weight.mat %*% input.activation
output.probability <<- output.activation/sum(output.activation)
mean.response <<- sum(output.layer * output.probability)
mean.response
}
```

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
#| code-overflow: wrap
alm.update <- function(corResp,c,lr){
  fz <- exp(-c*(output.layer - corResp)^2)
  teacherSignal <- (fz - output.activation)*lr
  #print(length(teacherSignal)); print(length(fz))
  wChange <- teacherSignal %*% t(input.activation)
  weight.mat <<- weight.mat + (wChange)
  weight.mat[weight.mat<0]=0 # prevent negative values
 # weight.mat[weight.mat>1]=1
  weight.mat <<- weight.mat
}

alm.trial <- function(input, corResp,c,lr){
  alm.response(input,c)
  alm.update(corResp,c,lr)
 # print(paste0("input=",input,"; corResp=",corResp,"; mean.response=",mean.response))
  mean.response
}
```

@@

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
exam.response <- function(input,c){
  # Find the index of the input node with the highest activation
  trainVec = sort(unique(xt))
  nearestTrain <- trainVec[which.min(abs(input - trainVec))]
  aresp <- alm.response(nearestTrain,c)
  #max.index <- which.max(input.activation)
  xUnder = ifelse(min(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) - 1])
  xOver = ifelse(max(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) + 1])

   exam.output = round(aresp + ((mOver - mUnder) / (xOver - xUnder)) * (input - nearestTrain), 3)
  # Determine the input nodes and associated weights for computing the slope
  exam.output
}
```

::: column-margin
###### Input node actvation

@@

### Prepare Simulation Data

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
# function to generate data, from either linear, exponential, quadratic or sinusoidal functions

generate.data <- function(x, type = "linear", noise = NA) {
  if (type == "linear") {
    y <- round(2.2*x + 30,0)
  }
  else if (type == "exponential") {
    y <- round(200*(1-exp(-x/25)),0)
  }
  else if (type == "sinusoidal") {
    y <- sin(2 * pi * x) 
  }
  else if (type == "quadratic") {
    y <- round(210 - ((x-50)^2)/12,0)
  }
  else {
    stop("type must be linear, exponential, quadratic, or sinusoidal")
  }
  # if noise is specified, add noise to the y values
  if(!is.na(noise)) {
    y <- y + round(rnorm(length(y), 0, noise),2)
  }
  data.frame(x, y,type)
}

envTypes <- c("linear", "exponential", "quadratic")

lowDensityTrainBlock <- c(30.5, 36.0, 41.0, 46.5, 53.5, 59.0, 64.0, 69.5)
medDensityTrainBlock <- c(
    30.0, 31.5, 33.0, 34.5, 36.5, 38.5, 41.0, 43.5, 46.0,
    48.5, 51.5, 54.0, 56.5, 59.0, 61.5, 63.5, 65.5, 67.0, 68.5, 70.0
)
highDensityTrainBlock <- c(
    30.0, 30.5, 31.0, 32.0, 33.0, 33.5, 34.5, 35.5,
    36.5, 37.0, 38.0, 38.5, 39.5, 40.5, 41.5, 42.0, 43.0,
    43.5, 44.5, 45.5, 46.5, 47.0, 48.0, 48.5, 49.0, 51.0, 51.5, 52.0,
    53.0, 53.5, 54.5, 55.5, 56.5, 57.0, 58.0, 58.5, 59.5, 60.5, 61.5, 
    62.0, 63.0,63.5, 64.5, 65.5, 66.5, 67.0, 68.0, 69.0, 69.5, 70.0
)
# all density conditions have the same # of training trials, but differ in the # of repetitions per items, or blocks,  low density has 25 training blocks, medium has 10 blocks, high has 4 blocks.

# generate training data, for each combination of environment type and density. Use purrr map functions. Rep each dataset by its number of blocks.
lowTrain <- map_dfr(envTypes, ~ generate.data(rep(lowDensityTrainBlock,25), type = .x)) %>% group_by(type) %>% mutate(block = rep(1:25, each = 8),trial=seq(1,200))

medTrain <- map_dfr(envTypes, ~ generate.data(rep(medDensityTrainBlock,10), type = .x)) %>% group_by(type) %>% mutate(block = rep(1:10, each = 20),trial=seq(1,200))

highTrain <- map_dfr(envTypes, ~ generate.data(rep(highDensityTrainBlock,4), type = .x)) %>% group_by(type) %>% mutate(block = rep(1:4, each = 50),trial=seq(1,200))

# nTrainExamples <- 10
# trainLowBound <-20; trainHIghBound <- 80
# nBlock <- 5 # number of times each training example is presented
# sample training data from function, nTrainExamples., between trainLowBound and trainHIghBound
# xt <- runif(nTrainExamples, trainLowBound, trainHIghBound)
# trainVec <- rep(xt, nBlock)

@@

```{r}
#| code-fold: show
#| 
# simulation function
alm.sim <- function(dat, c, lr,testRange=seq(0,100,.5)){

input.layer <<- matrix(seq(0,100,.5) ) # half step units for inputs, from 0 to 100
output.layer <<- matrix(seq(0,250,1)) # single step units for outputs, from 0 to 250
weight.mat <<- matrix(0.0000000,nrow=length(output.layer),ncol=length(input.layer )) # weights initialized to 0 (as in Delosh 1997)

xt<<-dat$x
# run training
st <- map2_dbl(dat$x, dat$y, ~alm.trial(.x,.y,c,lr))
# append training data to the data frame
dat <- dat %>% mutate(almResp = st)

return(list(d=dat,wm=weight.mat,c=c,lr=lr)) # final weightmat is probs incorrect for all but last
}

simOrganize <- function(simOut){
  dat <- simOut$d
  weight.mat <<- simOut$wm
  c <- simOut$c
  lr <- simOut$lr

  trainX <- unique(dat$x)
  xt <<- trainX

 almResp <- generate.data(seq(0,100,.5), type = first(dat$type)) %>% rowwise() %>% 
 mutate(model="ALM",resp = alm.response(x,c))

 examResp <- generate.data(seq(0,100,.5), type = first(dat$type)) %>% rowwise() %>% 
 mutate(model="EXAM",resp = exam.response(x,c))

 bind_rows(almResp,examResp) %>% 
 mutate(type=first(dat$type),
        c=c,lr=lr,
 type=factor(type,levels=c("linear","exponential","quadratic"))) %>%
 # compute test_region, equal to "train" if x is within trainX, interpolate if within trainX range, else "extrapolate"
  mutate(test_region = ifelse(x %in% trainX, "train", ifelse(x > min(trainX) & x < max(trainX), "interpolate", "extrapolate")))

}

@@

### Simulate Training:

```{r fig.width=10, fig.height=8}
#| column: page-inset-right
# split by type, then send each training dataset to simulation function
lowSim <- map(envTypes, ~ alm.sim(lowTrain %>% filter(type == .x), c = 1.4, lr = .4))

medSim <- map(envTypes, ~ alm.sim(medTrain %>% filter(type == .x), c = 1.4, lr = .4))

highSim <- map(envTypes, ~ alm.sim(highTrain %>% filter(type == .x), c = 1.4, lr = .4))

simAll %>% ggplot(aes(x=block,y=dev,color=type)) + stat_summary(geom="line",fun=mean,alpha=.4)+
  stat_summary(geom="point",fun=mean,alpha=.4)+
  stat_summary(geom="errorbar",fun.data=mean_cl_normal,alpha=.4)+facet_wrap(~density, scales="free_x")

@@

```{r fig.width=11,fig.height=10}
#| column: page-inset-right
#| 
lowSimTest <- map_dfr(lowSim,simOrganize) %>% mutate(density = "low")
medSimTest <- map_dfr(medSim,simOrganize) %>% mutate(density = "med")
highSimTest <- map_dfr(highSim,simOrganize) %>% mutate(density = "high")

simTestAll %>% ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  #geom_point(aes(x=x,y=y,color=test_region),alpha=.2)+ 
  geom_point(data=simTestAll %>% filter(test_region=="train"),aes(x=x,y=y),color="black",size=1,alpha=1) +
 # geom_point(data=simTestAll %>% filter(test_region %in% c("interpolate","extrapolate")),aes(x=x,y=y,color=test_region),alpha=.6) +
  # geom_point(data=simTestAll %>% filter(test_region=="extrapolate"),aes(x=x,y=y),color="purple",alpha=.3) +
  facet_grid(density~type) + 
  theme_bw() + theme(legend.position="bottom")

#lowSimTest %>% filter(model=="EXAM" & type=="linear")
#rm(list= ls()[sapply(ls(), function(x) class(get(x))) != 'function'])
#rm(weight.mat,input.activation,output.probability,output.activation,mean.response,xt)
# 
# simAll %>% ggplot(aes(x=block,y=dev)) + stat_summary(geom="line",fun=mean,alpha=.3)+stat_summary(geom="point",fun=mean)+
#   stat_summary(geom="errorbar",fun.data=mean_cl_normal)+facet_wrap(density~type, scales="free_x")

@@

@@

```{r}
#| eval: false
# Function that goes through every step of generating data, simulating training, and simulating generalization
full.sim <- function(c,lr,noise)
{

envTypes <- c("linear", "exponential", "quadratic")
lowDensityTrainBlock <- c(30.5, 36.0, 41.0, 46.5, 53.5, 59.0, 64.0, 69.5)
medDensityTrainBlock <- c(
    30.0, 31.5, 33.0, 34.5, 36.5, 38.5, 41.0, 43.5, 46.0,
    48.5, 51.5, 54.0, 56.5, 59.0, 61.5, 63.5, 65.5, 67.0, 68.5, 70.0
)
highDensityTrainBlock <- c(
    30.0, 30.5, 31.0, 32.0, 33.0, 33.5, 34.5, 35.5,
    36.5, 37.0, 38.0, 38.5, 39.5, 40.5, 41.5, 42.0, 43.0,
    43.5, 44.5, 45.5, 46.5, 47.0, 48.0, 48.5, 49.0, 51.0, 51.5, 52.0,
    53.0, 53.5, 54.5, 55.5, 56.5, 57.0, 58.0, 58.5, 59.5, 60.5, 61.5, 
    62.0, 63.0,63.5, 64.5, 65.5, 66.5, 67.0, 68.0, 69.0, 69.5, 70.0
)
# low density has 25 training blocks, medium has 10 blocks, high has 4 blocks. 
# generate training data, for each combination of environment type and density. Use purrr map functions. Rep each dataset by its number of blocks.
lowTrain <- map_dfr(envTypes, ~ generate.data(rep(lowDensityTrainBlock,25), type = .x, noise)) %>% group_by(type) %>% mutate(block = rep(1:25, each = 8),trial=seq(1,200))
medTrain <- map_dfr(envTypes, ~ generate.data(rep(medDensityTrainBlock,10), type = .x, noise)) %>% group_by(type) %>% mutate(block = rep(1:10, each = 20),trial=seq(1,200))
highTrain <- map_dfr(envTypes, ~ generate.data(rep(highDensityTrainBlock,4), type = .x, noise)) %>% group_by(type) %>% mutate(block = rep(1:4, each = 50),trial=seq(1,200))

lowSim <- map(envTypes, ~ alm.sim(lowTrain %>% filter(type == .x), c = 1.4, lr = .4))
medSim <- map(envTypes, ~ alm.sim(medTrain %>% filter(type == .x), c = 1.4, lr = .4))
highSim <- map(envTypes, ~ alm.sim(highTrain %>% filter(type == .x), c = 1.4, lr = .4))

@@

```{r fig.height=10,fig.width=11}
#| eval: false
k = full.sim(c=1.4,lr=.4,noise=2.0)
k4 = full.sim(c=1.4,lr=.4,noise=4.0)

@@

```{r}
#| eval: false

# label each each simulation with its density level (low, med, high), then combine all 3
lowSimTest <- lowSimTest %>% mutate(density = "low")
medSimTest <- medSimTest %>% mutate(density = "med")
highSimTest <- highSimTest %>% mutate(density = "high")

@@

# simOut <- map(c("lowTrain","medTrain","highTrain"), ~ map_dfr(envTypes, ~ alm.sim(get(.) %>% filter(type == .x), c = 1.4, lr = .4) %>% simOrganize))
# 
# simOut <- map_dfr(simGrid, ~ alm.sim(get(.x$trainData) %>% filter(type == .x$envTypes), c = 1.4, lr = .4) %>% simOrganize)
# 
# for each training dataset in c(lowTrain,medTrain,highTrain), run simulation function, separately for each type, and then bind rows of the output data frame.

@@

```{r}
#| code-fold: show
#| include: false
#| eval: false
alm.response <- function(input=1,c) {
input.activation <- exp(-c*(input.layer - input)^2)
input.activation <<- input.activation/sum(input.activation)
#print(length(input.activation)); print(dim(weight.mat))
output.activation <<- weight.mat %*% input.activation
output.probability <<- output.activation/sum(output.activation)
mean.response <<- sum(output.layer * output.probability)
mean.response
}

>>>>

```{r}
#lapply(c('tidyverse','data.table','igraph','ggraph','kableExtra'),library,character.only=TRUE))
pacman::p_load(tidyverse,data.table,igraph,ggraph,kableExtra, patchwork) 
purrr::walk(here::here(c("Functions/Display_Functions.R")),source)
source(here::here("Functions","deLosh_data.R"))
```

@@

ggplot() + geom_rect(data = odf,
            mapping = aes(xmin = xmin, ymin = ymin, 
                          xmax = xmax, ymax = ymax, 
                          fill = type, colour = type),alpha = 0.5) +
  geom_text(data=odf,aes(x=x,y=y,label=label,size=3)) +
  geom_path(data=plot_edges,mapping=aes(x=x,y=y,group=id,alpha=weight)) +
  theme_void() + theme(legend.position = "none")

```{r}

linear_function <- function(x) 2.2 * x + 30
exponential_function <- function(x) 200 * (1 - exp(-x/25))
quadratic_function <- function(x) 210 - (x - 50)^2 / 12

extrapLines <- list(geom_vline(xintercept=30,color="black",alpha=.4,linetype="dashed"),
                    geom_vline(xintercept=70,color="black",alpha=.4,linetype="dashed"))

linear_plot <- ggplot(deLosh_data$human_data_linear, aes(x, y)) +
    geom_point(shape=1) + stat_function(fun = linear_function, color = "black") +
  labs(y="Response Magnitude", title="Linear Function",x="") + extrapLines

exponential_plot <- ggplot(deLosh_data$human_data_exp, aes(x, y)) +
  geom_point(aes(shape = "Observed", color = "Observed"),shape=1) + 
  stat_function(aes(color = "True Function"),fun = exponential_function, geom="line")+
  labs(x="Stimulus Magnitude", title="Exponential Function",y="")  +
  extrapLines +
  scale_shape_manual(values = c(1)) +
  scale_color_manual(values = c("Observed" = "black", "True Function" = "black")) +
  theme(legend.title = element_blank(), legend.position="top") +
  guides(color = guide_legend(override.aes = list(shape = c(1, NA), 
                                                  linetype = c(0, 1))))

quadratic_plot <- ggplot(deLosh_data$human_data_quad, aes(x = x, y = y)) +
  geom_point( shape = 1) +
  stat_function( fun = quadratic_function, geom = "line") +
  labs(title="Quadratic Function",x="",y="") + extrapLines

linear_plot + exponential_plot + quadratic_plot

@@

@@

::: column-margin

###### Input node actvation

@@

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
alm.response <- function(input = 1, c, input.layer, output.layer,weight.mat) {
  input.activation <- exp(-c * (input.layer - input)^2) / sum(exp(-c * (input.layer - input)^2))
  output.activation <- weight.mat %*% input.activation
  output.probability <- output.activation / sum(output.activation)
  mean.response <- sum(output.layer * output.probability)
  list(mean.response = mean.response, input.activation = input.activation, output.activation = output.activation)
}
```

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
#| code-overflow: wrap
alm.update <- function(corResp, c, lr, output.layer, input.activation, output.activation, weight.mat) {
  fz <- exp(-c * (output.layer - corResp)^2)
  teacherSignal <- (fz - output.activation) * lr
  wChange <- teacherSignal %*% t(input.activation)
  weight.mat <- weight.mat + wChange
  weight.mat[weight.mat < 0] = 0
  return(weight.mat)
}

alm.trial <- function(input, corResp, c, lr, input.layer, output.layer, weight.mat) {
  alm_resp <- alm.response(input, c, input.layer,output.layer, weight.mat)
  updated_weight.mat <- alm.update(corResp, c, lr, output.layer, alm_resp$input.activation, alm_resp$output.activation, weight.mat)
  return(list(mean.response = alm_resp$mean.response, weight.mat = updated_weight.mat))
}
```

```{r}
#| code-fold: show
#| code-summary: "Toggle Code"
exam.response <- function(input, c, trainVec, input.layer = INPUT_LAYER_DEFAULT,output.layer = OUTPUT_LAYER_DEFAULT, weight.mat) {
  nearestTrain <- trainVec[which.min(abs(input - trainVec))]
  aresp <- alm.response(nearestTrain, c, input.layer = input.layer,output.layer = OUTPUT_LAYER_DEFAULT,weight.mat)$mean.response

  xUnder <- ifelse(min(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) - 1])
  xOver <- ifelse(max(trainVec) == nearestTrain, nearestTrain, trainVec[which(trainVec == nearestTrain) + 1])

  mUnder <- alm.response(xUnder, c, input.layer = input.layer, output.layer, weight.mat)$mean.response
  mOver <- alm.response(xOver, c, input.layer = input.layer,output.layer, weight.mat)$mean.response

  exam.output <- round(aresp + ((mOver - mUnder) / (xOver - xUnder)) * (input - nearestTrain), 3)
  exam.output
}
```

```{r}
#| code-fold: show

# simulation function
alm.sim <- function(dat, c, lr, input.layer = INPUT_LAYER_DEFAULT, output.layer = OUTPUT_LAYER_DEFAULT) {
  weight.mat <- matrix(0.00, nrow = length(output.layer), ncol = length(input.layer))
  xt <- dat$x
  n <- nrow(dat)
  st <- numeric(n) # Initialize the vector to store mean responses
  for(i in 1:n) {
    trial <- alm.trial(dat$x[i], dat$y[i], c, lr, input.layer, output.layer, weight.mat)
    weight.mat <- trial$weight.mat
    st[i] <- trial$mean.response
  }
  dat <- dat %>% mutate(almResp = st)
  return(list(d = dat, wm = weight.mat, c = c, lr = lr))
}

simOrganize <- function(simOut) {
  dat <- simOut$d
  weight.mat <- simOut$wm
  c <- simOut$c
  lr <- simOut$lr
  trainX <- unique(dat$x)

  almResp <- generate.data(seq(0,100,.5), type = first(dat$type)) %>% rowwise() %>% 
    mutate(model = "ALM", resp = alm.response(x, c, input.layer = INPUT_LAYER_DEFAULT,output.layer = OUTPUT_LAYER_DEFAULT, weight.mat = weight.mat)$mean.response)

  examResp <- generate.data(seq(0,100,.5), type = first(dat$type)) %>% rowwise() %>% 
    mutate(model = "EXAM", resp = exam.response(x, c, trainVec = trainX, input.layer = INPUT_LAYER_DEFAULT,output.layer = OUTPUT_LAYER_DEFAULT, weight.mat))

  organized_data <- bind_rows(almResp, examResp) %>% 
    mutate(type = first(dat$type),
           error = abs(resp - y),
           c = c,
           lr = lr,
           type = factor(type, levels = c("linear", "exponential", "quadratic")),
           test_region = ifelse(x %in% trainX, "train", 
                                ifelse(x > min(trainX) & x < max(trainX), "interpolate", "extrapolate")))
  organized_data
}

generateSimData <- function(density, envTypes, noise) {
  reps <- 200 / length(trainingBlocks[[density]])
  map_dfr(envTypes, ~ 
            generate.data(rep(trainingBlocks[[density]], reps), type = .x, noise)) |>
    group_by(type) |>
    mutate(block = rep(1:reps, each = length(trainingBlocks[[density]])),
           trial=seq(1,200))
}

simulateAll <- function(density,envTypes, noise, c = .2, lr = .2) {
  trainMat <- generateSimData(density, envTypes, noise)
  trainData <- map(envTypes, ~ alm.sim(trainMat %>% filter(type == .x), c = c, lr = lr))
  assign(paste(density),list(train=trainData, test=map_dfr(trainData, simOrganize) %>% mutate(density = density)))
}

@@

### Simulate Training and Testing

```{r fig.width=10, fig.height=8}
#| column: page-inset-right

envTypes <- c("linear", "exponential", "quadratic")
densities <- c("low", "med", "high")
noise=0
INPUT_LAYER_DEFAULT <- seq(0, 100, 0.5)
OUTPUT_LAYER_DEFAULT <- seq(0, 250, 1)

c = 1.4
lr=.8

results <- map(densities, ~ simulateAll(.x, envTypes, noise, c, lr)) |>
  set_names(densities)

trainAll <- results %>%
  map_df(~ map_df(.x$train, pluck, "d"), .id = "density") |>
  mutate(stage=as.numeric(cut(trial,breaks=20,labels=seq(1,20))),
         dev=sqrt((y-almResp)^2),
         density=factor(density,levels=c("low","med","high")),
         type=factor(type,levels=c("linear","exponential","quadratic"))) |>
  dplyr::relocate(density,type,stage)

simTestAll <- results |> map("test") |> bind_rows() |>
  group_by(type,density,model) %>%
  mutate(type=factor(type,levels=c("linear","exponential","quadratic")),
         density=factor(density,levels=c("low","med","high"))) %>%
  dplyr::relocate(density,type,test_region)

@@

### Training Data

```{r fig.width=10, fig.height=8}

trainAll %>% ggplot(aes(x=block,y=dev,color=type)) + stat_summary(geom="line",fun=mean,alpha=.4)+
  stat_summary(geom="point",fun=mean,alpha=.4)+
  stat_summary(geom="errorbar",fun.data=mean_cl_normal,alpha=.4)+facet_wrap(~density, scales="free_x")

@@

```{r fig.width=11,fig.height=10}
#| column: page-inset-right
#| 
simTestAll %>% ggplot(aes(x=x,y=y)) + 
  geom_point(aes(x=x,y=resp,shape=model,color=model),alpha=.7,size=1) + 
  geom_line(aes(x=x,y=y),alpha=.4)+ 
  geom_point(data=simTestAll %>% filter(test_region=="train"),aes(x=x,y=y),color="black",size=1,alpha=1) +
  facet_grid(density~type) + 
  theme_bw() + theme(legend.position="bottom")

@@

### Table

| Model & Definition                    | R Code                                                                                                      |
|---------------------------------------|-------------------------------------------------------------------------------------------------------------|
| $a_i(X)=\exp \left|-\gamma \cdot\left[X-X_i\right]^2\right|$   | `exp(-c * (input.layer - input)^2)`        |
| $o_j(X)=\Sigma_{i=1, M} w_{j i} \cdot a_i(X)$                 | `weight.mat %*% input.activation`                                                     |
| $P\left[Y_j \mid X\right]=o_j(X) / \Sigma_{k=1, L} o_k(X)$   | `output.activation / sum(output.activation)`                                         |
| $m(X)=\Sigma_{j=1, L} Y_j \cdot P\left[Y_j \mid X\right]$    | `sum(output.layer * output.probability)`                                                  |
| $f_j(Z)=e^{-c\cdot(Z-Y_j)^2}$                                 | `exp(-c * (output.layer - corResp)^2)`                                                               |
| $w_{ji}(t+1)=w_{ji}(t)+\alpha \cdot {f_i(Z(t))-O_j(X(t))} \cdot a_i(X(t)$   | ` lr *(fz - output.activation) %*% t(input.activation)`  |
                                                      |
| $E[Y|X_i]=m(X_i) + \bigg[\frac{m(X_{i+1})-m(X_{i-1})}{X_{i+1} - X_{i-1}} \bigg]\cdot[X-X_i]$ | ` trainVec[which.min(abs(input - trainVec))]; xUnder <- ...; xOver <- ...; mUnder <- ...; mOver <- ...; exam.output <- round(aresp + ((mOver - mUnder) / (xOver - xUnder)) * (input - nearestTrain), 3)`|

@@

```{r}
#| eval: false
#| include: false 
# Function that goes through every step of generating data, simulating training, and simulating generalization
full.sim <- function(c,lr,noise)
{

envTypes <- c("linear", "exponential", "quadratic")
lowDensityTrainBlock <- c(30.5, 36.0, 41.0, 46.5, 53.5, 59.0, 64.0, 69.5)
medDensityTrainBlock <- c(
    30.0, 31.5, 33.0, 34.5, 36.5, 38.5, 41.0, 43.5, 46.0,
    48.5, 51.5, 54.0, 56.5, 59.0, 61.5, 63.5, 65.5, 67.0, 68.5, 70.0
)
highDensityTrainBlock <- c(
    30.0, 30.5, 31.0, 32.0, 33.0, 33.5, 34.5, 35.5,
    36.5, 37.0, 38.0, 38.5, 39.5, 40.5, 41.5, 42.0, 43.0,
    43.5, 44.5, 45.5, 46.5, 47.0, 48.0, 48.5, 49.0, 51.0, 51.5, 52.0,
    53.0, 53.5, 54.5, 55.5, 56.5, 57.0, 58.0, 58.5, 59.5, 60.5, 61.5, 
    62.0, 63.0,63.5, 64.5, 65.5, 66.5, 67.0, 68.0, 69.0, 69.5, 70.0
)
# low density has 25 training blocks, medium has 10 blocks, high has 4 blocks. 
# generate training data, for each combination of environment type and density. Use purrr map functions. Rep each dataset by its number of blocks.
a
lowSim <- map(envTypes, ~ alm.sim(lowTrain %>% filter(type == .x), c = 1.4, lr = .4))
medSim <- map(envTypes, ~ alm.sim(medTrain %>% filter(type == .x), c = 1.4, lr = .4))
highSim <- map(envTypes, ~ alm.sim(highTrain %>% filter(type == .x), c = 1.4, lr = .4))

@@

```{r fig.height=10,fig.width=11}
#| eval: false
#| include: false
k = full.sim(c=1.4,lr=.4,noise=2.0)
k4 = full.sim(c=1.4,lr=.4,noise=4.0)

@@

```{r}
#| eval: false
#| include: false
# label each each simulation with its density level (low, med, high), then combine all 3
lowSimTest <- lowSimTest %>% mutate(density = "low")
medSimTest <- medSimTest %>% mutate(density = "med")
highSimTest <- highSimTest %>% mutate(density = "high")

@@

# simOut <- map(c("lowTrain","medTrain","highTrain"), ~ map_dfr(envTypes, ~ alm.sim(get(.) %>% filter(type == .x), c = 1.4, lr = .4) %>% a))
# 
# simOut <- map_dfr(simGrid, ~ alm.sim(get(.x$trainData) %>% filter(type == .x$envTypes), c = 1.4, lr = .4) %>% simOrganize)
# 
# for each training dataset in c(lowTrain,medTrain,highTrain), run simulation function, separately for each type, and then bind rows of the output data frame.

@@

```{r}
#| include: false
#| eval: false
alm.response <- function(input=1,c) {
input.activation <- exp(-c*(input.layer - input)^2)
input.activation <<- input.activation/sum(input.activation)
#print(length(input.activation)); print(dim(weight.mat))
output.activation <<- weight.mat %*% input.activation
output.probability <<- output.activation/sum(output.activation)
mean.response <<- sum(output.layer * output.probability)
mean.response
}
